Sparse Function-space Representation \\ of Neural Networks

### 2023-05-14 paper/main.tex
rewrite 2.1 first draft done

%
\subsection{Linearization gives rise to a Gaussian process} 
\label{sec:nn2gp}
%
Quantifying uncertainty in a Bayesian framework boils down to finding a posterior of $p(\vw \mid \dataset)$, which represents the uncertainty over our parameters based on the prior and the data. The posterior structure is related to the loss-landscape around the MAP weights $\vw^*$, which induces a correlation structure on functions produced from the neural network's deterministic mapping. Samples from the weight posterior are perturbed versions of $f_{\vw^*}$ characterizing uncertainty in the learned model in the function space. Although the true posterior is intractable given the non-linearities of the neural network, sampling techniques available, such as Hamiltonian Monte-Carlo (HMC), in most cases, are not available given the computational burden.
 The uncertainty plotted \cref{fig:banana} shows the induced function space posterior by sampling through the model.

A common approach is to approximate the correlation structure of a distribution centred at the MAP estimate as done in the Laplace-GGN \citep{khan2019approximate, daxberger2021laplace, maddox2021fast}. The Laplace GGN takes the MAP solution and approximates the Hessian of our loss function  $\ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i)$ with the GGN approximation. Another way of viewing the approximation is that we build the following approximate model $\hat{f}_\weights(\mathbf{x}_{i}) = \Jac{\weights}{\mathbf{x}_i} \vw $  where $\Jac{\weights}{\mathbf{x}} \coloneqq \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$ is the Jacobian as was pointed out in \citep{khan2019approximate}. Using the Hessian of the approximate model, we arrive at the Laplace-GGN posterior for $q(\vw)$.

In this paper we are concerned with learning posterior directly in the function space that is having a $q(\vf)$ instead of a $q(\vw)$. A common choice is to use a Gaussian process characterizes a distribution over functions in terms of the first two moments, a mean $\mu(\cdot)$ and covariance function $\kappa(\cdot,\cdot)$ (kernel). We aim to find a kernel and mean function based on the neural network to build the Gaussian process posterior.  A linear function in terms of parameters can be converted from the weight space to the function space (see Ch.~XXX in \cite{rasmussen2006gaussian} as follows:
\begin{equation} \label{eq:weight_func}
g_\weights(\mathbf{x}_{i}) = \phi^\top\!(\vx) \, \vw \implies \mu(\vx) = \vzeros \quad \kappa(\mathbf{x}, \mathbf{x}') = \frac{1}{\delta} \phi(\vx)^\top \phi(\vx)
\end{equation}

- distribution over w given to you by the non-linear function. One could sample over weights pass it over the function and you have distribution over functions. HMC is heavy, structure of the problem around the map estimate. characteruze this as the first two moments of the distributions over the functions restrict to the first two moments.

Replace with a taylor series at the MAP first order taylor expansion in w. degenerate cov function.

- more citations there.

- GP induced by the trained neural network. not GP prior.

- Add the NTK bit back in.

- assume that when you are at the map estimate when you have converged this should be small.

- Write out that w beomces zero under the expectation.

-  transition from background to methods.

- More explination of the dual equation 5 the loss landscape link. Perbutation point linked directly to that here.

-

%A Gaussian process characterizes a distribution over functions in terms of the first two moments. The key quantity of a Gaussian process is the prior covariance function (or kernel function) $\kappa(\vx,\vx')$, where $\vx, \vx'$ are input vectors in $\R^{\mathrm{D}}$. We aim to find a kernel based on the neural network to build the Gaussian process posterior.  A linear function in terms of parameters can easily be converted from the weight space to the function space as follows:
%\begin{equation}
%g_\weights(\mathbf{x}_{i})) = \phi(\vx)^\top \vw \implies \mu(\vx) = \vzeros \quad \kappa(\mathbf{x}, \mathbf{x}') = \frac{1}{\delta} \phi(\vx)^\top \phi(\vx)
%\end{equation}
%
%
%
% The approximation can be seen as taking the map solution of \cref{eq-empirical-risk} as the mean of $q(\vw)$ but then substituting the following functional approximation for $\hat{f}_\weights(\mathbf{x}_{i}) = \Jac{\weights^*}{\mathbf{x}} $

%A Gaussian procss characterizes a distribution over functions in terms of the first two moments. The key quantity of 
%A Gaussian process is the prior covariance function (or kernel function) $\kappa(\vx,\vx')$, where $\vx, \vx'$ are input vectors in $\R^{\mathrm{D}}$. We aim to find a kernel based on the neural network to build the Gaussian process posterior. 
%A GP is fully characterized by a mean $\mu(\cdot)$ and covariance function $\kappa(\cdot,\cdot)$ (kernel).
%If we had the following function parameterization $g(\vx) = \phi(\vx)^\top \vw$ where $\phi(\vx)$ is a feature map, we can convert from weight space prior to function space prior giving the following kernel $\kappa(\vx, \vx') = \delta^{-1} \phi(\vx) \phi^\T(\vx')$ (see Ch.~XXX in \cite{rasmussen2006gaussian} for details).\todo{this part needs to be rewritten, starts talking about a general function parameterisation and then jumps directly to the neural network} 
%\subsection{Neural Network to GP Prior} 
%\label{sec:nn2gp}
%
A Gaussian process characterizes a distribution over functions in terms of the first two moments, a mean $\mu(\cdot)$ and covariance function $\kappa(\cdot,\cdot)$ (kernel). We aim to find a kernel and mean function based on the neural network to build the Gaussian process posterior.  A linear function in terms of parameters can be converted from the weight space to the function space (see Ch.~XXX in \cite{rasmussen2006gaussian} as follows:
\begin{equation} \label{eq:weight_func}
g_\weights(\mathbf{x}) = \phi^\top\!(\vx) \, \vw \implies \mu(\vx) = \vzeros \quad \kappa(\mathbf{x}, \mathbf{x}') = \frac{1}{\delta} \phi(\vx)^\top \phi(\vx)
\end{equation}
Given an approximate posterior $q(\vw)$, \citet{immerImprovingPredictionsBayesian2021} proposed to make the following linear approximation
$f^{\textrm{lin.}}_{\vw^*}(\vx; \vw) = f_{\vw^*}(\vx) + \Jac{\weights^*}{\mathbf{x}} \, (\vw - \vw^*)$ of the neural network, where $\Jac{\weights}{\mathbf{x}} \coloneqq \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$ is the Jacobian. The approximation can be used with any approximate inference technique but in the paper they focus on the Laplace-GGN. They find that the linear approximation helps to improve predictions by correcting the approximate posterior when sampled through a non-Gaussian likelihoods.

Furthermore, under the linear approximation they showed one can convert the weight space prior to the function space by combining with \cref{eq:weight_func} meaning the converged neural network GP prior is:
\begin{equation} \label{eq-laplace-approx-function-space}
% g(\vx) \sim \GP \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad \text{with} \quad
  \mu(\vx) =  f^{\textrm{lin.}}_{\vw_*}(\vx; \vw = \vzeros) = f_{\vw^*}(\vx) - \Jac{\weights^*}{\mathbf{x}} \vw^* \quad \text{and} \quad
  \kappa(\mathbf{x}, \mathbf{x}')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}}\JacT{\weights^*}{\mathbf{x}'},
\end{equation}
%
%where the kernel is the so-called Neural Tangent Kernel (NTK, \cite{jacot2018neural}) with Jacobian $\Jac{\weights}{\mathbf{x}} \coloneqq \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$. Alternatively, we could have followed the derivation by \citet{immerImprovingPredictionsBayesian2021}, where they perform a local linearization of the neural network  $f^{\textrm{lin.}}_{\vw^*}(\vx; \vw) = f_{\vw^*}(\vx) + \Jac{\weights^*}{\mathbf{x}} \, (\vw - \vw^*)$. Their mean function differ from ours, \cref{eq-laplace-approx-function-space}, as theirs comes from setting  $\mu(\vx) = f^{\textrm{lin.}}_{\vw_*}(\vx; \vw = \vzeros) = f_{\vw^*}(\vx) - \Jac{\weights^*}{\mathbf{x}} \vw^*$, \ie the difference between the linear model approximation and the neural network function at $\vx$. %which does not seem useful as a prior mean function, 
%On the other hand, we set $\mu(\vx) = \vzeros$, as common in GP literature and derived from a feature map conversion. \todo{derived what?}

\section{Methods}
\paragraph{Linearization in the dual parameters}
As stated in \citet[Lemma~1]{csato2002sparse}, the posterior process mean and covariance functions of a GP prior, combined with a likelihood function and $\dataset$ can be stated as:
\begin{equation}  \label{eq:gp_pred}
 \myexpect_{p(f_i \mid\vy)}[f_i] = \vk_{\vx i}^\top \valpha \quad \text{and} \quad
\mathrm{Var}_{p(f\mid\vy)}[f_i] = \kappa_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},
\end{equation}
where $f_i = f_\weights(\vx_i)$, the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{m \times m}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $\kappa_{ii} = \kappa(\vx_i, \vx_i)$.  The dual parameters consist of the vector $\valpha \in \R^{m}$ and vector $\vbeta \in \R^{m}$, defined as:
\begin{equation}
\label{eq:dual_param}
\alpha_i \coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f_i}\log p(y_i \mid f_i)],  \quad
\beta_i \coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f_i f_i}\log p(y_i \mid f_i)] ,
\end{equation}
%
The parameters $(\alpha_i, \beta_i)$ are the dual parameters that arise in the dual formulations used in, for example, support vector machines \citep{cortes1995support}, whose origins are found in work by \citet{kimeldorf1971some}. The above relations holds for generic likelihoods, and no approximations are involved as the expectation is under the exact posterior. The question is how do we form the dual variables without access to the exact posterior and different approximate inference techniques solve the question in different ways. Bound techniques such as \citet{jaakkola1997variational} attempt to bound the integral thus making the integral tractable, while Expecation Propogation (EP) as used in \citet{csato2002sparse} iteratively finds dual variables but sets them to log of the expectation of the approximate posterior. Meanwhile variational inference  \cite{khan2017conjugate, adam2021dual} find them using the variational approximation of the posterior to solve the integral.

>>>>

%
\subsection{Linearization gives rise to a Gaussian process} 
\label{sec:nn2gp}
%

Quantifying uncertainty in a Bayesian framework means finding a posterior of $p(\vw \mid \dataset)$. The posterior represents the uncertainty over our parameters based on the prior and the data. The posterior structure directly relates to the optimization loss around the MAP weights $\vw^*$. Furthermore, the landscape in the weight space induces functions produced by the neural network's deterministic mapping. If one was to sample from the weight posterior, the corresponding functions created can be viewed as perturbed versions of the $f_{\vw^*}$. In most applications, we care about predictions from the neural network, so the distribution in the function space not in the weight space.

Although the true posterior is intractable given the non-linearities of the neural network sampling techniques are available such as Hamiltonian Monte-Carlo (HMC), The uncertainty plotted \cref{fig:banana} shows the induced function space posterior by sampling through the model. However, in most cases, sampling techniques like HMC are not available, given the computational burden.

A common approach is to approximate the correlation structure of a distribution centred at the MAP estimate as done in the Laplace-GGN \citep{khan2019approximate, daxberger2021laplace, maddox2021fast}. The Laplace GGN takes the MAP solution and approximates the Hessian of the loss function  $\ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i)$ with the GGN approximation. Another way of viewing the approximation is that we build the following approximate model $\hat{f}_\weights(\mathbf{x}_{i}) = \Jac{\weights_*}{\mathbf{x}_i} \vw_i $  where $\Jac{\weights}{\mathbf{x}} \coloneqq \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$ is the Jacobian as was pointed out in \citep{khan2019approximate}. Using the Hessian of the approximate model, we arrive at the Laplace-GGN posterior for $q(\vw)$. As pointed out in \citet{khan2019approximate}, a Gaussian posterior covariance approximation in the weight space implies a covariance structure in the function space.

In this paper we are concerned with learning posterior directly in the function space that is having a $q(\vf)$ instead of a $q(\vw)$. We choose is to use a Gaussian process an extension of the multivariate normal that characterizes a distribution over functions in terms of the first two moments, a mean $\mu(\cdot)$ and covariance function $\kappa(\cdot,\cdot)$ (kernel). We aim to find a kernel and mean function based on the neural network to build the Gaussian process posterior.  A linear function in terms of parameters can be converted from the weight space to the function space (see Ch.~XXX in \cite{rasmussen2006gaussian} as follows:
\begin{equation} \label{eq:weight_func}
g_\weights(\mathbf{x}_{i}) = \phi^\top\!(\vx) \, \vw \implies \mu(\vx) = 0 \quad \kappa(\mathbf{x}, \mathbf{x}') = \frac{1}{\delta} \phi(\vx)^\top \phi(\vx)
\end{equation}

We can therefore use the approximate linear model defined by the GGN approximation to define our GP derived from the trained neural network,
\begin{equation} \label{eq-laplace-approx-function-space}
% g(\vx) \sim \GP \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad \text{with} \quad
  \mu(\vx) =  0 \quad \text{and} \quad
  \kappa(\mathbf{x}, \mathbf{x}')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}}\JacT{\weights^*}{\mathbf{x}'},
\end{equation}
where the kernel is the so-called Neural Tangent Kernel (NTK, \cite{jacot2018neural}), given we have a kernel function, we can then combine it with the $\dataset$ to form the posterior. In \citep{khan2019approximate} they took a similar approach but instead of fitting the GP posterior to the actual data they adjusted the data to be pseudo data. Similarly, using a first order approximation of the neural network function to form a GLM model \citet{immerImprovingPredictionsBayesian2021} also arrives at the same NTK function for the GP derived from the neural network, however a different mean function. Both previous approaches have attempted to adjust the GP posterior mean function to ensure the predictions are from the neural network $f_{\vw^*}(\vx)$. In our method, we wish to make predictions from the derived GP and thus do not make such adjustments. We now present the general formulation of the functional representation of our trained neural network

\section{Methods}
\paragraph{Linearization in the dual parameters}
As stated in \citet[Lemma~1]{csato2002sparse}, the posterior process mean and covariance functions of a GP prior, combined with a likelihood function and $\dataset$ can be stated as:
\begin{equation}  \label{eq:gp_pred}
 \myexpect_{p(f_i \mid\vy)}[f_i] = \vk_{\vx i}^\top \valpha \quad \text{and} \quad
\mathrm{Var}_{p(f\mid\vy)}[f_i] = \kappa_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},
\end{equation}
where $f_i = f_\weights(\vx_i)$, the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{m \times m}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $\kappa_{ii} = \kappa(\vx_i, \vx_i)$.  The dual parameters consist of the vector $\valpha \in \R^{m}$ and vector $\vbeta \in \R^{m}$, defined as:
\begin{equation}
\label{eq:dual_param}
\alpha_i \coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f_i}\log p(y_i \mid f_i)],  \quad
\beta_i \coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f_i f_i}\log p(y_i \mid f_i)] ,
\end{equation}\todo{ More explination of the dual equation 6 the loss landscape link. Perbutation point linked directly to that here. }
%
The parameters $(\alpha_i, \beta_i)$ are the dual parameters that arise in the dual formulations used in, for example, support vector machines \citep{cortes1995support}, whose origins are found in work by \citet{kimeldorf1971some}. The above relations holds for generic likelihoods, and no approximations are involved as the expectation is under the exact posterior. The question is how do we form the dual variables without access to the exact posterior and different approximate inference techniques solve the question in different ways. Bound techniques such as \citet{jaakkola1997variational} attempt to bound the integral thus making the integral tractable, while Expecation Propogation (EP) as used in \citet{csato2002sparse} iteratively finds dual variables but sets them to log of the expectation of the approximate posterior. Meanwhile variational inference  \cite{khan2017conjugate, adam2021dual} find them using the variational approximation of the posterior to solve the integral.
